The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-04 paper/ref_approach.tex
some edits + bodyfat vs projpred analysis new without fixing variables

\begin{document}
\maketitle
\begin{abstract}
When a reference model is used to guide the variable selection problem, it acts as a noise-filter. We show how this translate into higher stability and goodness of the selection. We devise a simple reference model approach that can be used on top of any feature selection procedure and use the normal means problem as a benchmark comparing different methods using or not the reference model. We include in our comparisons state-of-the-art complete selection procedures showing improved stability, false discovery rate and sensitivity of the selection, and also Bayesian methods controlling the amount of posterior shrinkage and bias. In addition to that we show in a real-world example the benefits achieved by the reference model through the projection predictive approach compared to the selection via stepwise backward regression.
\end{abstract}

@@

As a motivational example to the study of the consequences of the use of a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples in the submodels. The R-package \textit{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Assume to have $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of S posterior samples of the vector parameter of the model.
\item Rank the covariates using a heuristics and consider as candidate submodels the nested ones, starting to include the best ranked feature. The submodel are naturally identify by their model size.
\item For each submodel, project the reference model samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL}\;[p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\,||\,p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})] \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the model distribution of the reference model with parameters fixed at the sample value $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariates related to the statistical unit (identified by the subscript $i$ in the variable $\tilde{y}_{i}$), whereas $p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ the same but for the submodel. The projected samples correspond to the submodel parameter distribution.
\item For each submodel (size), test the predictive performance.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model prediction utility score.
\end{enumerate}

Expression \eqref{eq:draw_by_draw} is not an easy optimisation problem, but in the special case of submodels in the GLMs family, \eqref{eq:draw_by_draw} reduces to a maximum likelihood estimation problem, which can be easily optimised (e.g. using iteratively reweighted least squares alogirthm, IRLS). For further details on \textit{projpred} workflow and implementation see \cite{paper:projpred}.

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{johnson1996fitting}. The analysis is mainly taken from Vehtari's R-notebook available at \url{https://avehtari.github.io/modelselection/bodyfat.html}, adjusting only the number of bootstrap repetitions to 1000 to have a consistent comparison with results from \cite{paper:bodyfat}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure, and the original covariates are 13 anthropometric measurements (e.g. height, weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection, for a total of 251 observations. The goal is to find the simplest model which is able to describe, that is to predict, sufficiently well the amount of fat.

\cite{paper:bodyfat} report the results using the stepwise backward elimination with a significance level of 0.157 with AIC selection, fixing abdomen and height to be always included in the model. We implement the selection via projection: the reference model includes all the covariates using as a prior the regularised horseshoe, submodels are explored using a forward search and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. In order to automatise the procedure, we select the submodel size as the smallest one which has an elpd score higher than the reference model score with probability 0.05. Figure \ref{fig:inclusion_frequencies} shows the bootstrap inclusion frequencies combining the results from \cite{paper:bodyfat} for the stepwise backward elimination and our experiments for the projection. The projection predictive approach has only two variables with inclusion frequencies above 50\% (abdomen is the only included always), the third most included is `wrist' at 49.5\% and the fourth is `height' at 27.8\%, while the stepwise regression has seven features above 50\%, of which abdomen and height are fixed. Such a lower stability of the stepwise regression can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. The first five selected models have a cumulative frequency of almost 70\% with projpred, whereas with the stepwise regression only of 11.8\%. In addition to that we note that the size of the selected model is much smaller when using projpred.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 1000 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l||l|r||l|r}
  \hline
M & projpred & Freq (\%) & steplm & Freq (\%)  \\ 
  \hline
1 & abdomen, weight & 34.2 & abdomen, height, wrist, age, chest, biceps & 3.2 \\
2 & abdomen, wrist & 13.3 & abdomen, height, wrist, age, neck, forearm, thigh, hip & 2.9 \\
3 & abdomen, weight, wrist & 7.6 & abdomen, height, wrist, age, forearm, chest & 1.9 \\
4 & abdomen, height & 7.4 & abdomen, height, wrist, age, neck, forearm, chest & 1.9 \\
5 & abdomen, height, wrist & 6.7 & abdomen, height, wrist, age, neck, forearm, chest, thigh, hip & 1.9 \\
6 & abdomen, age, wrist & 2.9 & abdomen, height, wrist, age, neck, chest, biceps & 1.8 \\
7 & abdomen, height, neck & 2.2 & abdomen, height, wrist, age, neck, thigh, biceps, hip & 1.6 \\
8 & abdomen, age, height, wrist & 1.2 & abdomen, height, wrist, age, neck, forearm & 1.5 \\
9 & abdomen, weight, thigh & 1.1 & abdomen, height, wrist, age, neck, biceps & 1.5 \\
10 & abdomen, weight, neck & 1.0 & abdomen, height, wrist, age, beck, forearm, chest, biceps & 1.4 \\
   \hline
\end{tabular}
\caption{Bootstrap model selection frequencies after 1000 bootstrap samples.}
\label{tab:model_frequencies}
\end{table}

Table \ref{tab:model_performances} reports the predictive performances, in terms of root mean square error, of the full model and the correspondent selected model using both the projection predictive approach and the stepwise backward elimination. We also repeat the selection after adding uncorrelated noisy features up to a total of 100 covariates and report in such a case both the size of the selected model and the number of selected noisy features. Whilst with the original data we do not observe any gain using projpred in terms of predictive performance (we do observe a smaller size for the selected model, however with a much higher computational cost), when noisy features are added the projection is able to notably improve the predictive performance keeping the same submodel size, whereas the stepwise regression selects a much larger submodel including a large number (32 over 38) of noisy features.

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l|r|r|r|r}
  \hline
 & rmse.full & rmse.sel & size.sel & noisy.sel \\ 
  \hline
projpred & 4.4 & 4.5 & 2 & -  \\
steplm & 4.4 & 4.5 & 7 & - \\
\hline
\hline
projpred & $^{*}$4.4 & $^{*}$4.5 & 2 & 0  \\
steplm & 5.3 & 5.5 & 38 & 32 \\
   \hline
\end{tabular}
\caption{Model predictive performances with original data (first two rows) and adding noisy features (last two rows). Results with an asterisk using 10-fold CV, otherwise 20-fold CV.}
\label{tab:model_performances}
\end{table}

@@

Figure \ref{fig:k} shows the posterior means for the shrinkage factors of each parameter using the regularised horseshoe prior, averaged after 100 data simulations. Relevant and non-relevant features are respectively highlighted in red and blue. The ideal shrinkage would have $\kappa=0$ when the parameter is a signal, whereas $\kappa=1$ when it is noise. Respectively on the y-axis and on the x-axis are depicted the shrinkage factors using the reference model approach or not. If the reference model did not make any difference, we would expect the points to lie on the diagonal. Non-relevant features happen to lie on average on the diagonal for all the simulated scenario, meaning not evident benefits coming from the reference model. Although when considering the shrinkage amount for the relevant ones, we note that the points always lie under the diagonal, meaning that the reference model is able to shrink less the true signals. Such benefit is proportional to the difficulty of the problem, that is it increases with lower correlation and number of observations. Note that the perfect distinction of the two clusters of features (relevant and non-relevant) observable on both the marginal distributions of the shrinkage factors is due to the averaging over the 100 data simulations. In a single data realisation there would be more noise and the clusters would be closer and overlapping.
The results for the Dirichlet-Laplace prior are very similar and they are reported in the Appendix \ref{ap:shrinkage_signal}.

@@

\hypertarget{complete-selection}{%
\subsubsection{Complete selection analysis}\label{complete-selection}}
As complete selection procedures we consider the control of the local false discovery rate \citep{paper:efron, efron2012large}, the empirical Bayes median \citep{johnstone2004needles} and the selection by posterior credible intervals. 
\\
The control of the local false discovery rate is applied through the R-package \textit{locfdr}. It consists in testing the z-values $\{z_{j}\}_{j=1}^{p}$ to belong to the theoretical null distribution $f_{0}$ (the null hypothesis, meaning no relevance) against the alternative hypothesis distribution $f_{1}$. In our case $f_{0}$ correspond to the standard normal distribution, see expression \eqref{eq:normal_means_problem2}. The quantity of interest is the local false discovery rate (loc.fdr) defined as:
\
\begin{equation}
loc.fdr(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)}
\end{equation}
where $\pi_{0}$ is the prior probability of the null hypothesis and $f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal distribution of the z-values. The latter is estimated using splines with 7 degrees of freedom and we select features with local false discovery rate under the value 0.2. Such a value is commonly suggested only because it corresponds to a Bayes factor large than 36 (assuming $\pi_{0}\geq0.9$), other values can be considered and in our experience do not affect the results of the comparison. We use the default setting to estimate $\pi_{0}$ from the data provided by the package.
\\
The empirical Bayes median procedure is given by the R-package \textit{EbayesThresh} and consists in fitting a Bayesian model with a prior composed by a mixture of a delta in zero and a heavy-tailed distribution. We use, as suggested, a Laplace distribution resulting in a thresholding property, i.e. there exists a threshold value such that all the data under that threshold have posterior median equal to zero. Therefore the selection is done including those parameters whose posterior median is different from zero. The hyperparameter of the Laplace distribution and the mixing weight of the prior are estimated by marginal maximum likelihood. 
\\
The selection by posterior credible intervals is done using the regularised horseshoe prior and selecting those features whose posterior distribution does not include zero in the interval between the 0.05-th and the 0.95-th quantiles.

@@

We conclude our experiments using the real world dataset body fat, already described in the comparison between the projection predictive approach and the stepwise regression in Section \ref{projection}. We carry out two different analyses: one using the normal means problem framework and the other one using the stepwise backward regression. In the former, we add to the original data noisy uncorrelated covariates up to a total of 1000 features. These covariates are normally distributed and scaled as the original ones. We compute correlations between each variable and the target variable, that is the amount of fat, and transform them by Fisher transformation. The original assumption, in order that \eqref{eq:fisher_transformation} holds, is that the variables are jointly normally distributed. In our experience the normal approximation in \eqref{eq:fisher_transformation} is still reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the variance to be one, instead we estimate it from the data. The methods we compare are those used in Section \ref{complete-selection}: the control of the local false discovery rate (loc.fdr), the empirical Bayes median (EB.med) and the selection by posterior credible intervals at level 0.9 (ci.90). The maximum likelihood estimator is used to estimate the variance of the null hypothesis in the local false discovery rate \cite[see][Chap. 6]{efron2012large}, whereas the median absolute deviation from zero for the empirical Bayes median method. The selection via posterior credible intervals is done using a regularised horseshoe prior on the parameters for the means, while a standard log-normal  prior for the variance. In order to vary the difficulty of the selection, we bootstrap subsamples of different sizes, going from $n=50$ up to $n=251$. Each time results are averaged over 100 bootstrap subsamples.

@@

In the second analysis, we repeat the selection of Section \ref{projection} implementing the stepwise backward regression by AIC as implemented in the \textit{step} function in R. We have not fixed any variable to be always included in the model and the overall number of covariates (original plus noisy) is 100, as it was in the last part of Section \ref{projection}. We compare results using or not the reference model on top of the procedure; here we have used again the reference model given by the first five supervised principal components. Figure \ref{fig:bodyfat_step_refvsdata} shows the number of noisy features included in the final model and the out-of-sample root mean square error (RMSE). Results are after 100 bootstrap samples on the whole dataset, and the predictive performance is tested on the observations excluded at each bootstrap sample. We observe that the reference model reduces the number of noisy features included in the final model. This leads to less overfit to the data and, thus, improved out-of-sample predictive performance in terms of RMSE. Also in this case, we see that there is not a notable difference in terms of the stability of the two approaches, which in this case is qualitatively assessed through the variability of the histogram. We can not use the stability measure provided by \cite{paper:stability}, since it does not take in account possibly swaps in the set of selected features due to high correlation between the covariates, which in this case can occur. Our reference model approach applied to the stepwise backward regression achieves outstanding improvements considering its simplicity, yet it does not reach the goodness of more thoughtful implementations, as projpred (see results of Section \ref{projection}).

@@

This paper has discussed the importance of using a reference model when dealing with feature selection, or more in general model reduction. We want to highlight that the goal of this paper is not to provide a specific algorithm or procedure to carry out the selection, but rather to bring motivations and attentions to the general use of a reference model. 
\\
The reference model acts as an approximation of the data generation mechanism through its predictive distribution. Such approximation is generally less noisy than the sample estimation available from the observed data, leading to the main benefits of the reference model approach. In our comparisons, we have analysed the effect of a reference model in the form of a filter on the observed target values on top of different variable selection methods widely used. The overall benefits we have showed independently of the specific method applied, consisting in stability and goodness of the selection, motivate what we refer to as the reference model approach for the model reduction framework. Such approach translates into a family of different methods, some of which have been present in the literature for several years.
\\
In a real-world data application (body fat data), we have applied one of these methods, namely the projection predictive approach as indicated by \cite{paper:projpred}, in opposition to the selection via stepwise backward regression with the results provided by \cite{paper:bodyfat}. The selection via projection resulted to be more stable and led to a sparser representation. After adding additional noisy covariates, the selected submodel showed also higher predictive performance compared to the one selected by the stepwise backward regression, which included several non-relevant features. Applying a reference model on top of the stepwise backward regression resulted in a minor number of non-relevant features selected and an improved predictive performance, yet without reaching projection predictive approach results.
\\
Whenever it is possible to come up with a reasonable reference model, our suggestion is to employ it when selecting models, because of its nature as baseline of the comparisons and its ability to improve the selection. Note that one of the main challenges in many real world application will consist in devising the reference model itself and assessing its predictive performance, for which we suggest the user to rely on robust modelling workflow indications \cite[see e.g.][]{gabry2019visualization}.
\\
All Bayesian models have been implemented using Stan \citep{paper:stan}, which uses HMC-NUTS sampler \citep{hoffman2014no} and allows full posterior Bayesian inference. The code to run all the experiments is available at \url{https://github.com/fpavone/ref-approach-paper}.

@@

\appendix

\hypertarget{ap:shrinkage_signal}{%
\section{Appendix: Shrinkage factors with Dirichlet-Laplace prior}\label{ap:shrinkage_signal}}
Figure \ref{fig:k_DL} shows the posterior mean shrinkage factors using the Dirichlet-Laplace prior for the normal means problem in the simulated data study of Section \ref{comparison}. The result is qualitatively equal to the regularised horseshoe prior's one (see Figure \ref{fig:k}).

>>>>

\begin{document}
\maketitle
\begin{abstract}
When a reference model is used to guide the variable selection problem, it acts as a noise-filter modelling the data generation mechanism. We show how this translate into higher stability and goodness of the selection. We devise a simple reference model approach that can be used on top of any feature selection procedure and use the normal means problem as a benchmark comparing different methods using or not the reference model. We include in our comparisons state-of-the-art complete selection procedures showing improved stability, false discovery rate and sensitivity of the selection, and also Bayesian methods controlling the amount of posterior shrinkage and bias. In addition to that we show in a real-world example the benefits achieved by the reference model through the projection predictive approach compared to the selection via stepwise backward regression.
\end{abstract}

@@

As a motivational example to the study of the consequences of the use of a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples in the submodels. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Assume to have $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of S posterior samples of the vector parameter of the model.
\item Rank the covariates using a heuristics and consider as candidate submodels the nested ones, starting to include the best ranked feature. The submodel are naturally identify by their model size.
\item For each submodel, project the reference model samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL}\;[p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\,||\,p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})] \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the model distribution of the reference model with parameters fixed at the sample value $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariates related to the statistical unit (identified by the subscript $i$ in the variable $\tilde{y}_{i}$), whereas $p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ the same but for the submodel. The projected samples correspond to the submodel parameter distribution.
\item For each submodel (size), test the predictive performance.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model prediction utility score.
\end{enumerate}

Expression \eqref{eq:draw_by_draw} is not an easy optimisation problem, but in the special case of submodels in the GLMs family, \eqref{eq:draw_by_draw} reduces to a maximum likelihood estimation problem, which can be easily optimised (e.g. using iteratively reweighted least squares alogirthm, IRLS). For further details on \texttt{projpred} workflow and implementation see \cite{paper:projpred}.

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{johnson1996fitting}. The analysis is based on from Vehtari's R-notebook available at \url{https://avehtari.github.io/modelselection/bodyfat.html}, which is inspired by \cite{paper:bodyfat}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure, and the original covariates are 13 anthropometric measurements (e.g. height, weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection, for a total of 251 observations. The goal is to find the simplest model which is able to describe, that is to predict, sufficiently well the amount of fat. We use the same data again later in Section \ref{real-world-data}.

\cite{paper:bodyfat} report results using the stepwise backward elimination with a significance level of 0.157 with AIC selection, fixing abdomen and height to be always included in the model. We implement the stepwise backward elimination via AIC without fixing any variable, and the selection via projection: the reference model includes all the covariates using as a prior the regularised horseshoe, submodels are explored using a forward search and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in \texttt{R}: the stepwise elimination is carried out combining the \texttt{step} and \texttt{lm} functions, whereas the selection via projection using the \texttt{projpred} package. In order to automatise the latter, we select the submodel size as the smallest one which has an elpd score higher than the reference model score with probability 0.05. The inclusion frequencies of each feature in the final model after 100 bootstrap samples are depicted in Figure \ref{fig:inclusion_frequencies}: the projection predictive approach has only two variables with inclusion frequencies above 50\% (abdomen is the only included always), the third most included is `wrist' at 44\% and the fourth is `height' at 35\%, while the stepwise regression has seven features above 50\%. Such a lower stability of the stepwise regression can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. The first five selected models have a cumulative frequency of 76\% with projpred, whereas with the stepwise regression only of 14\%. In addition to that we note that the size of the selected model is much smaller when using projpred.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 100 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l||l|r||l|r}
  \hline
M & projpred & Freq (\%) & steplm & Freq (\%)  \\ 
  \hline
1 & abdomen, weight & 39 & abdomen, age, forearm, height, hip, neck, thigh, wrist & 4 \\
2 & abdomen, wrist & 10 & abdomen, age, chest, forearm, height, neck, thigh, wrist & 4 \\
3 & abdomen, height & 10 & abdomen, forearm, height, neck, wrist & 2 \\
4 & abdomen, height, wrist & 9 & abdomen, forearm, neck, weight, wrist & 2 \\
5 & abdomen, weight, wrist & 8 & abdomen, age, height, hip, thigh, wrist & 2 \\
6 & abdomen, chest, height, wrist & 2 & abdomen, age, height, hip, neck, thigh, wrist & 2 \\
7 & abdomen, biceps, weight, wrist & 2 & abdomen, age, ankle, forearm, height, hip, neck, thigh, wrist & 2 \\
8 & abdomen, height, weight, wrist & 2 & abdomen, age, biceps, chest, height, neck, wrist & 2 \\
9 & abdomen, age, wrist & 2 & abdomen, age, biceps, chest, forearm, height, neck, thigh, wrist & 2 \\
10 & abdomen, age, height, neck, thigh, wrist & 2 & abdomen, age, ankle, biceps, weight, wrist & 2 \\
   \hline
\end{tabular}
\caption{Bootstrap model selection frequencies after 100 bootstrap samples.}
\label{tab:model_frequencies}
\end{table}

%\begin{table}[tp]
%\scriptsize
%\centering
%\begin{tabular}{l||l|r||l|r}
%  \hline
%M & projpred & Freq (\%) & steplm & Freq (\%)  \\ 
%  \hline
%1 & abdomen, weight & 34.2 & abdomen, height, wrist, age, chest, biceps & 3.2 \\
%2 & abdomen, wrist & 13.3 & abdomen, height, wrist, age, neck, forearm, thigh, hip & 2.9 \\
%3 & abdomen, weight, wrist & 7.6 & abdomen, height, wrist, age, forearm, chest & 1.9 \\
%4 & abdomen, height & 7.4 & abdomen, height, wrist, age, neck, forearm, chest & 1.9 \\
%5 & abdomen, height, wrist & 6.7 & abdomen, height, wrist, age, neck, forearm, chest, thigh, hip & 1.9 \\
%6 & abdomen, age, wrist & 2.9 & abdomen, height, wrist, age, neck, chest, biceps & 1.8 \\
%7 & abdomen, height, neck & 2.2 & abdomen, height, wrist, age, neck, thigh, biceps, hip & 1.6 \\
%8 & abdomen, age, height, wrist & 1.2 & abdomen, height, wrist, age, neck, forearm & 1.5 \\
%9 & abdomen, weight, thigh & 1.1 & abdomen, height, wrist, age, neck, biceps & 1.5 \\
%10 & abdomen, weight, neck & 1.0 & abdomen, height, wrist, age, beck, forearm, chest, biceps & 1.4 \\
%   \hline
%\end{tabular}
%\caption{Bootstrap model selection frequencies after 1000 bootstrap samples.}
%\label{tab:model_frequencies}
%\end{table}

Table \ref{tab:model_performances} reports the predictive performances, in terms of root mean square error, of the full model and the correspondent selected model using both the projection predictive approach and the stepwise backward elimination. We also repeat the selection after adding uncorrelated noisy features up to a total of 100 covariates and report in such a case both the size of the selected model and the number of selected noisy features. Whilst with the original data we do not observe any gain using projpred in terms of predictive performance (we do observe a smaller size for the selected model, however with a much higher computational cost), when noisy features are added the projection is able to notably improve the predictive performance keeping the same submodel size, whereas the stepwise regression selects a much larger submodel including a large number (13 over 19) of noisy features.

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l|r|r|r|r}
  \hline
 & rmse.full & rmse.sel & size.sel & noisy.sel \\ 
  \hline
projpred & 4.35 & 4.47 & 2 & -  \\
steplm & 4.38 & 4.47 & 7 & - \\
\hline
\hline
projpred & 4.58 & 4.54 & 2 & 0  \\
steplm & 5.89 & 5.43 & 19 & 13 \\
   \hline
\end{tabular}
\caption{Model predictive performances with original data (first two rows) and adding noisy features (last two rows) estimated with 10-fold cross-validation.}
\label{tab:model_performances}
\end{table}

%\begin{table}[tp]
%\scriptsize
%\centering
%\begin{tabular}{l|r|r|r|r}
%  \hline
% & rmse.full & rmse.sel & size.sel & noisy.sel \\ 
%  \hline
%projpred & 4.4 & 4.5 & 2 & -  \\
%steplm & 4.4 & 4.5 & 7 & - \\
%\hline
%\hline
%projpred & $^{*}$4.4 & $^{*}$4.5 & 2 & 0  \\
%steplm & 5.3 & 5.5 & 38 & 32 \\
%   \hline
%\end{tabular}
%\caption{Model predictive performances with original data (first two rows) and adding noisy features (last two rows). Results with an asterisk using 10-fold CV, otherwise 20-fold CV.}
%\label{tab:model_performances}
%\end{table}

@@

Figure \ref{fig:k} shows the posterior means for the shrinkage factors of each parameter using the regularised horseshoe prior, averaged after 100 data simulations. Relevant and non-relevant features are respectively highlighted in red and blue. The ideal shrinkage would have $\kappa=0$ when the parameter is a signal, whereas $\kappa=1$ when it is noise. Respectively on the y-axis and on the x-axis are depicted the shrinkage factors using the reference model approach or not. If the reference model did not make any difference, we would expect the points to lie on the diagonal. Non-relevant features happen to lie on average on the diagonal for all the simulated scenario, meaning not evident benefits coming from the reference model. Although when considering the shrinkage amount for the relevant ones, we note that the points always lie under the diagonal, meaning that the reference model is able to shrink less the true signals. Such benefit is proportional to the difficulty of the problem, that is it increases with lower correlation and number of observations. Note that the perfect distinction of the two clusters of features (relevant and non-relevant) observable on both the marginal distributions of the shrinkage factors is due to the averaging over the 100 data simulations. In a single data realisation there would be more noise and the clusters would be closer and overlapping.
The results for the Dirichlet-Laplace prior are very similar and we do not include them in the paper.

@@

\hypertarget{complete-selection}{%
\subsubsection{Complete selection analysis}\label{complete-selection}}
As complete selection procedures we consider the control of the local false discovery rate \citep{paper:efron, efron2012large}, the empirical Bayes median \citep{johnstone2004needles} and the selection by posterior credible intervals. 
\\
The control of the local false discovery rate is applied through the \texttt{R}-package \texttt{locfdr}. It consists in testing the z-values $\{z_{j}\}_{j=1}^{p}$ to belong to the theoretical null distribution $f_{0}$ (the null hypothesis, meaning no relevance) against the alternative hypothesis distribution $f_{1}$. In our case $f_{0}$ correspond to the standard normal distribution, see expression \eqref{eq:normal_means_problem2}. The quantity of interest is the local false discovery rate (loc.fdr) defined as:
\
\begin{equation}
loc.fdr(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)}
\end{equation}
where $\pi_{0}$ is the prior probability of the null hypothesis and $f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal distribution of the z-values. The latter is estimated using splines with 7 degrees of freedom and we select features with local false discovery rate under the value 0.2. Such a value is commonly suggested only because it corresponds to a Bayes factor large than 36 (assuming $\pi_{0}\geq0.9$), other values can be considered and in our experience do not affect the results of the comparison. We use the default setting to estimate $\pi_{0}$ from the data provided by the package.
\\
The empirical Bayes median procedure is given by the \texttt{R}-package \texttt{EbayesThresh} and consists in fitting a Bayesian model with a prior composed by a mixture of a delta in zero and a heavy-tailed distribution. We use, as suggested, a Laplace distribution resulting in a thresholding property, i.e. there exists a threshold value such that all the data under that threshold have posterior median equal to zero. Therefore the selection is done including those parameters whose posterior median is different from zero. The hyperparameter of the Laplace distribution and the mixing weight of the prior are estimated by marginal maximum likelihood. 
\\
The selection by posterior credible intervals is done using the regularised horseshoe prior and selecting those features whose posterior distribution does not include zero in the interval between the 0.05-th and the 0.95-th quantiles.

@@

We conclude our experiments using one more time the real world dataset body fat, already described in the comparison between the projection predictive approach and the stepwise regression in Section \ref{projection}. We carry out two different analyses: one using the normal means problem framework and the other one using the stepwise backward regression. In the former, we add to the original data noisy uncorrelated covariates up to a total of 1000 features. These covariates are normally distributed and scaled as the original ones. We compute correlations between each variable and the target variable, that is the amount of fat, and transform them by Fisher transformation. The original assumption, in order that \eqref{eq:fisher_transformation} holds, is that the variables are jointly normally distributed. In our experience the normal approximation in \eqref{eq:fisher_transformation} is still reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the variance to be one, instead we estimate it from the data. The methods we compare are those used in Section \ref{complete-selection}: the control of the local false discovery rate (loc.fdr), the empirical Bayes median (EB.med) and the selection by posterior credible intervals at level 0.9 (ci.90). The maximum likelihood estimator is used to estimate the variance of the null hypothesis in the local false discovery rate \cite[see][Chap. 6]{efron2012large}, whereas the median absolute deviation from zero for the empirical Bayes median method. The selection via posterior credible intervals is done using a regularised horseshoe prior on the parameters for the means, while a standard log-normal  prior for the variance. In order to vary the difficulty of the selection, we bootstrap subsamples of different sizes, going from $n=50$ up to $n=251$. Each time results are averaged over 100 bootstrap subsamples.

@@

In the second analysis, we repeat the selection of Section \ref{projection} via stepwise backward regression by AIC. In this case the overall number of covariates (original plus noisy) is 100, as it was in the last part of Section \ref{projection}. We compare results using or not the reference model on top of the procedure according to our simple reference model approach; here the reference model is again given by the first five supervised principal components. Figure \ref{fig:bodyfat_step_refvsdata} shows the number of noisy features included in the final model and the out-of-sample root mean square error (RMSE). Results are after 100 bootstrap samples on the whole dataset, and the predictive performance is tested on the observations excluded at each bootstrap sample. We observe that the reference model reduces the number of noisy features included in the final model. This leads to less overfit to the data and, thus, improved out-of-sample predictive performance in terms of RMSE. Also in this case, we see that there is not a notable difference in terms of the stability of the two approaches, which in this case is qualitatively assessed through the variability of the histogram. We can not use the stability measure provided by \cite{paper:stability}, since it does not take in account possibly swaps in the set of selected features due to high correlation between the covariates, which in this case can occur. Our reference model approach applied to the stepwise backward regression achieves outstanding improvements considering its simplicity, yet it does not reach the goodness of more thoughtful implementations, as projpred (see results of Section \ref{projection}).

@@

This paper has discussed the importance of using a reference model when dealing with feature selection, or more in general model reduction. We want to highlight that the goal of this paper is not to provide a specific algorithm or procedure to carry out the selection, but rather to bring motivations and attentions to the general use of a reference model. 
\\
The reference model acts as an approximation of the data generation mechanism through its predictive distribution. Such approximation is generally less noisy than the sample estimation available from the observed data, leading to the main benefits of the reference model approach. In our comparisons, we have analysed the effect of a reference model in the form of a filter on the observed target values on top of different variable selection methods widely used. The overall benefits we have showed independently of the specific method applied, consisting in stability and goodness of the selection, motivate what we refer to as the reference model approach for the model reduction framework. Such approach translates into a family of different methods, some of which have been present in the literature for several years.
\\
In a real-world data application (body fat data), we have applied one of these methods, namely the projection predictive approach as indicated by \cite{paper:projpred}, in opposition to the selection via stepwise backward regression with the results provided by \cite{paper:bodyfat}. The selection via projection resulted to be more stable and led to a sparser representation. After adding additional noisy covariates, the selected submodel showed also higher predictive performance compared to the one selected by the stepwise backward regression, which included several non-relevant features. Applying a reference model on top of the stepwise backward regression resulted in a minor number of non-relevant features selected and an improved predictive performance, yet without reaching the projection predictive approach results.
\\
Whenever it is possible to come up with a reasonable reference model, our suggestion is to employ it when selecting models, because of its nature as baseline of the comparisons and its ability to improve the selection. Note that one of the main challenges in many real world application will consist in devising the reference model itself and assessing its predictive performance, for which we suggest the user to rely on robust modelling workflow indications \cite[see e.g.][]{gabry2019visualization}.
\\
All Bayesian models have been implemented using \texttt{Stan} \citep{paper:stan}, which uses HMC-NUTS sampler \citep{hoffman2014no} and allows full posterior Bayesian inference. The code to run all the experiments is available at \url{https://github.com/fpavone/ref-approach-paper}.

@@

%\appendix

%\hypertarget{ap:shrinkage_signal}{%
%\section{Appendix: Shrinkage factors with Dirichlet-Laplace prior}\label{ap:shrinkage_signal}}
%Figure \ref{fig:k_DL} shows the posterior mean shrinkage factors using the Dirichlet-Laplace prior for the normal means problem in the simulated data study of Section \ref{comparison}. The result is qualitatively equal to the regularised horseshoe prior's one (see Figure \ref{fig:k}).

%\begin{figure}[tp]
%  \centering
%  \includegraphics[width=0.98\textwidth]{graphics/k_DL.pdf}
%  \caption{Average posterior mean values for the shrinkage factors using the Dirichlet-Laplace prior after 100 simulations. Respectively in red and in blue relevant and non-relevant variables.\\}
%  \label{fig:k_DL}
%\end{figure}
